Observing and Predicting Knowledge Worker Stress, Focus and Awakeness in the Wild

### 2020-10-01 IJHCSSubmission/elsarticle-template.tex
reviewed Intro

\cortext[ma]{Main Author}

\begin{abstract}
Knowledge workers face many challenges in the workplace: work is 
fragmented, disruptions are constant, tasks are complex, and work hours can be 
long. These challenges can impact knowledge workers' stress, focus and awakeness, and in turn their interaction with the digital environment, the quality of work 
performed and their productivity in general. We report on a field study with 14
knowledge workers over an eight-week period in which we investigated, using
experience sampling, how the workers experience stress \rev{and} awakeness over time. During this field study, we also collected biometric data including heart- and skin-related measures, which we then used to investigate if it is possible to predict stress, focus and awakeness, in the moment.
We observed and report on various trends in knowledge worker stress and awakeness levels
over several weeks, finding that people tend to have certain ``baseline'' levels for these aspects and stressful days or days with low awakeness tend to cluster. We further show that machine learning models can be built
from the data of a single minimally invasive sensor
to
predict stress, focus and awakeness. Overall, we found that our models were capable of large improvements in precision and recall in comparison to a random classifier for stress (25.9\% increase over random for precision, 4.2\% for recall) and awakeness (52.4\% increase in precision, 40.8\% in recall). The abstract concept of focus proved to be
the hardest to predict (26.0\% increase in precision, 27.8\% decrease in recall).
\end{abstract}

>>>>

\cortext[ma]{Main Authors}

\begin{abstract}
Knowledge workers face many challenges in the workplace: work is 
fragmented, disruptions are constant, tasks are complex, and work hours can be 
long. These challenges can impact knowledge workers' stress, focus and awakeness, and in turn their interaction with the digital environment, the quality of work 
performed and their productivity in general. We report on a field study with 14
knowledge workers over an eight-week period in which we investigated, using
experience sampling, how the workers experience stress \rev{and} awakeness over time. During this field study, we also collected biometric data including heart- and skin-related measures, which we then used to investigate if it is possible to predict stress, focus and awakeness, in the moment.
We observed and report on various trends in knowledge worker stress and awakeness levels
over several weeks, finding that people tend to have certain ``baseline'' levels for these aspects\rev{. Moreover, we found that days with high levels of stress tend to cluster, similarly as the days with low awakeness}. We further show that machine learning models can be built
from the data of a single minimally invasive \rev{device} to
predict stress, focus, and awakeness. Overall, we found that our models were capable of large improvements in precision and recall in comparison to a random classifier for stress (25.9\% increase over random for precision, 4.2\% for recall) and awakeness (52.4\% increase in precision, 40.8\% in recall). The abstract concept of focus proved to be the hardest to predict (26.0\% increase in precision, 27.8\% decrease in recall).
\end{abstract}
